---
title: "CUT&Tag Analysis Report"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: bootstrap
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(DT)
library(tidyverse)
library(viridis)
library(GenomicRanges)
library(chromVAR)

# Read input data
samplesheet <- read_tsv(snakemake@params[["samplesheet"]])
stats <- read_tsv(snakemake@input[["stats"]])
seacr_num <- read_tsv(snakemake@input[["seacr_num"]])
seacr_reprod <- read_tsv(snakemake@input[["seacr_reprod"]])
seacr_frip <- read_tsv(snakemake@input[["seacr_frip"]])
seacr_width <- read_tsv(snakemake@input[["seacr_width"]])
macs3_num <- read_tsv(snakemake@input[["macs3_num"]])
macs3_reprod <- read_tsv(snakemake@input[["macs3_reprod"]])
macs3_frip <- read_tsv(snakemake@input[["macs3_frip"]])
macs3_width <- read_tsv(snakemake@input[["macs3_width"]])
consolidated <- read_tsv(snakemake@input[["consolidated"]])

# Read TSS enrichment if available
tss_enrichment <- NULL
if (length(snakemake@input[["tss_enrichment"]]) > 0 &&
    file.exists(snakemake@input[["tss_enrichment"]])) {
  tss_enrichment <- read_tsv(snakemake@input[["tss_enrichment"]])
}
```

Overview {data-icon="fa-chart-line"}
=======================================================================

Column {data-width=350}
-----------------------------------------------------------------------

### Sample Summary

```{r}
summary_stats <- consolidated %>%
  select(sampleID, condition, seacr_peakN, macs3_peakN, seacr_FRiP, macs3_FRiP) %>%
  mutate(
    seacr_peakN = round(seacr_peakN, 0),
    macs3_peakN = round(macs3_peakN, 0),
    seacr_FRiP = round(seacr_FRiP, 2),
    macs3_FRiP = round(macs3_FRiP, 2)
  )

DT::datatable(summary_stats,
              options = list(pageLength = 10, autoWidth = TRUE),
              colnames = c("Sample", "Condition", "SEACR Peaks", "MACS3 Peaks",
                          "SEACR FRiP", "MACS3 FRiP"),
              caption = "Summary of peak calling results for all samples")
```

### Alignment Statistics

```{r}
DT::datatable(stats,
              options = list(pageLength = 10, autoWidth = TRUE),
              caption = "Alignment and quality control statistics")
```

Column {data-width=650}
-----------------------------------------------------------------------

### Peak Numbers by Method and Condition

```{r}
peak_comparison <- consolidated %>%
  select(sampleID, condition, seacr_peakN, macs3_peakN) %>%
  pivot_longer(cols = c(seacr_peakN, macs3_peakN),
               names_to = "method", values_to = "peak_count") %>%
  mutate(method = case_when(
    method == "seacr_peakN" ~ "SEACR",
    method == "macs3_peakN" ~ "MACS3"
  ))

p1 <- peak_comparison %>%
  ggplot(aes(x = condition, y = peak_count, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  geom_point(aes(color = method), position = position_jitterdodge(dodge.width = 0.75)) +
  scale_fill_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  scale_color_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  theme_minimal() +
  labs(title = "Peak Count Comparison",
       x = "Condition", y = "Number of Peaks",
       fill = "Method", color = "Method")

ggplotly(p1)
```

### FRiP Scores by Method and Condition

```{r}
frip_comparison <- consolidated %>%
  select(sampleID, condition, seacr_FRiP, macs3_FRiP) %>%
  pivot_longer(cols = c(seacr_FRiP, macs3_FRiP),
               names_to = "method", values_to = "frip_score") %>%
  mutate(method = case_when(
    method == "seacr_FRiP" ~ "SEACR",
    method == "macs3_FRiP" ~ "MACS3"
  ))

p2 <- frip_comparison %>%
  ggplot(aes(x = condition, y = frip_score, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  geom_point(aes(color = method), position = position_jitterdodge(dodge.width = 0.75)) +
  scale_fill_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  scale_color_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  theme_minimal() +
  labs(title = "FRiP Score Comparison",
       x = "Condition", y = "FRiP Score (%)",
       fill = "Method", color = "Method")

ggplotly(p2)
```

Peak Analysis {data-icon="fa-mountain"}
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

### Peak Width Distribution - SEACR

```{r}
# Reconstruct peak width data for SEACR
suffix_seacr <- ".peaks.stringent.bed"
peakW_seacr <- tibble()

for (absname in snakemake@input[["seacr_peaks"]]) {
  if (file.size(absname) > 0) {
    filename <- str_split(absname, "/")[[1]][3]
    sample <- str_split(filename, suffix_seacr)[[1]][1]
    peakInfo <- read.table(absname, header = FALSE, fill = TRUE) %>%
      mutate(width = abs(V3 - V2))
    peakW_seacr <- tibble(width = peakInfo$width, sampleID = sample) %>%
      bind_rows(peakW_seacr)
  }
}

if (nrow(peakW_seacr) > 0) {
  df_seacr <- left_join(peakW_seacr, samplesheet) %>%
    select(sampleID, condition, width)

  p3 <- df_seacr %>%
    ggplot(aes(x = width, fill = condition)) +
    geom_histogram(bins = 50, alpha = 0.7) +
    scale_fill_viridis_d(begin = 0.1, end = 0.8, option = "magma") +
    scale_x_log10() +
    facet_wrap(~condition, scales = "free_y") +
    theme_minimal() +
    labs(title = "SEACR Peak Width Distribution",
         x = "Peak Width (bp)", y = "Count")

  ggplotly(p3)
}
```

### Peak Width Distribution - MACS3

```{r}
# Reconstruct peak width data for MACS3
suffix_macs <- "_peaks.narrowPeak"
peakW_macs <- tibble()

for (absname in snakemake@input[["macs3_peaks"]]) {
  if (file.size(absname) > 0) {
    filename <- str_split(absname, "/")[[1]][3]
    sample <- str_split(filename, suffix_macs)[[1]][1]
    peakInfo <- read.table(absname, header = FALSE, fill = TRUE) %>%
      mutate(width = abs(V3 - V2))
    peakW_macs <- tibble(width = peakInfo$width, sampleID = sample) %>%
      bind_rows(peakW_macs)
  }
}

if (nrow(peakW_macs) > 0) {
  df_macs <- left_join(peakW_macs, samplesheet) %>%
    select(sampleID, condition, width)

  p4 <- df_macs %>%
    ggplot(aes(x = width, fill = condition)) +
    geom_histogram(bins = 50, alpha = 0.7) +
    scale_fill_viridis_d(begin = 0.1, end = 0.8, option = "magma") +
    scale_x_log10() +
    facet_wrap(~condition, scales = "free_y") +
    theme_minimal() +
    labs(title = "MACS3 Peak Width Distribution",
         x = "Peak Width (bp)", y = "Count")

  ggplotly(p4)
}
```

Column {data-width=500}
-----------------------------------------------------------------------

### Peak Width Statistics

```{r}
width_stats_combined <- bind_rows(
  seacr_width %>% mutate(method = "SEACR"),
  macs3_width %>% mutate(method = "MACS3")
) %>%
  select(sampleID, condition, method, median_width, mean_width, pct_narrow, pct_broad)

DT::datatable(width_stats_combined,
              options = list(pageLength = 15, autoWidth = TRUE),
              colnames = c("Sample", "Condition", "Method", "Median Width",
                          "Mean Width", "% Narrow", "% Broad"),
              caption = "Peak width statistics by sample and method") %>%
  formatRound(columns = c("median_width", "mean_width"), digits = 1) %>%
  formatRound(columns = c("pct_narrow", "pct_broad"), digits = 2)
```

### Peak Categories Comparison

```{r}
peak_categories <- width_stats_combined %>%
  select(sampleID, condition, method, pct_narrow, pct_broad) %>%
  pivot_longer(cols = c(pct_narrow, pct_broad),
               names_to = "category", values_to = "percentage") %>%
  mutate(category = case_when(
    category == "pct_narrow" ~ "Narrow (≤500bp)",
    category == "pct_broad" ~ "Broad (>2000bp)"
  ))

p5 <- peak_categories %>%
  ggplot(aes(x = sampleID, y = percentage, fill = category)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  scale_fill_viridis_d(begin = 0.2, end = 0.8, option = "plasma") +
  facet_wrap(~method, scales = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Peak Category Distribution",
       x = "Sample", y = "Percentage (%)",
       fill = "Peak Category")

ggplotly(p5, height = 500)
```

Quality Control {data-icon="fa-check-circle"}
=======================================================================

Column {data-width=600}
-----------------------------------------------------------------------

### Reproducibility Analysis

```{r}
reprod_combined <- bind_rows(
  seacr_reprod %>% mutate(method = "SEACR"),
  macs3_reprod %>% mutate(method = "MACS3")
)

p6 <- reprod_combined %>%
  ggplot(aes(x = sampleID, y = peakReprodRate, fill = condition)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  scale_fill_viridis_d(begin = 0.1, end = 0.8, option = "magma") +
  facet_wrap(~method, scales = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Peak Reproducibility Rate",
       x = "Sample", y = "Reproducibility Rate (%)",
       fill = "Condition")

ggplotly(p6)
```

### Fragment Length Distribution

```{r}
# Read fragment length data
fragment_files <- snakemake@input[["fragments"]]
fragment_data <- tibble()

for (file in fragment_files) {
  if (file.exists(file)) {
    sample <- str_extract(basename(file), "^[^.]+")
    frags <- read_tsv(file, col_names = c("length", "count"))
    fragment_data <- bind_rows(fragment_data,
                              frags %>% mutate(sampleID = sample))
  }
}

if (nrow(fragment_data) > 0) {
  fragment_data <- fragment_data %>%
    left_join(samplesheet, by = "sampleID") %>%
    filter(length <= 1000)  # Focus on relevant range

  p7 <- fragment_data %>%
    ggplot(aes(x = length, y = count, color = condition)) +
    geom_line(alpha = 0.7) +
    scale_color_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
    facet_wrap(~sampleID, scales = "free_y") +
    theme_minimal() +
    labs(title = "Fragment Length Distribution",
         x = "Fragment Length (bp)", y = "Count",
         color = "Condition")

  ggplotly(p7)
}
```

Column {data-width=400}
-----------------------------------------------------------------------

### TSS Enrichment Scores {data-height=400}

```{r}
if (!is.null(tss_enrichment)) {
  tss_with_condition <- tss_enrichment %>%
    left_join(samplesheet, by = c("Sample" = "sampleID"))

  p8 <- tss_with_condition %>%
    ggplot(aes(x = condition, y = TSS_Enrichment, fill = condition)) +
    geom_boxplot(alpha = 0.7) +
    geom_jitter(width = 0.2) +
    scale_fill_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
    theme_minimal() +
    labs(title = "TSS Enrichment Scores",
         x = "Condition", y = "TSS Enrichment Score",
         fill = "Condition")

  ggplotly(p8)
} else {
  htmltools::div(
    style = "text-align: center; padding: 50px;",
    htmltools::h4("TSS enrichment analysis not performed"),
    htmltools::p("Add a GTF file to config.json to enable TSS enrichment calculation")
  )
}
```

### Quality Metrics Summary {data-height=600}

```{r}
qc_metrics <- consolidated %>%
  select(sampleID, condition, seacr_FRiP, macs3_FRiP,
         seacr_median_width, macs3_median_width,
         seacr_pct_narrow, macs3_pct_narrow)

if (!is.null(tss_enrichment)) {
  qc_metrics <- qc_metrics %>%
    left_join(tss_enrichment, by = c("sampleID" = "Sample"))
}

DT::datatable(qc_metrics,
              options = list(pageLength = 10, autoWidth = TRUE, scrollX = TRUE),
              caption = "Comprehensive quality control metrics") %>%
  formatRound(columns = c("seacr_FRiP", "macs3_FRiP"), digits = 2) %>%
  formatRound(columns = c("seacr_median_width", "macs3_median_width"), digits = 1) %>%
  formatRound(columns = c("seacr_pct_narrow", "macs3_pct_narrow"), digits = 2)
```

Method Comparison {data-icon="fa-balance-scale"}
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### SEACR vs MACS3 Peak Counts

```{r}
p9 <- consolidated %>%
  ggplot(aes(x = seacr_peakN, y = macs3_peakN, color = condition)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
  scale_color_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  theme_minimal() +
  labs(title = "SEACR vs MACS3 Peak Count Correlation",
       x = "SEACR Peak Count", y = "MACS3 Peak Count",
       color = "Condition")

ggplotly(p9)
```

### SEACR vs MACS3 FRiP Scores

```{r}
p10 <- consolidated %>%
  ggplot(aes(x = seacr_FRiP, y = macs3_FRiP, color = condition)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
  scale_color_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  theme_minimal() +
  labs(title = "SEACR vs MACS3 FRiP Score Correlation",
       x = "SEACR FRiP Score (%)", y = "MACS3 FRiP Score (%)",
       color = "Condition")

ggplotly(p10)
```

### Peak Width Comparison

```{r}
p11 <- consolidated %>%
  ggplot(aes(x = seacr_median_width, y = macs3_median_width, color = condition)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
  scale_color_viridis_d(begin = 0.1, end = 0.8, option = "plasma") +
  theme_minimal() +
  labs(title = "SEACR vs MACS3 Median Peak Width",
       x = "SEACR Median Width (bp)", y = "MACS3 Median Width (bp)",
       color = "Condition")

ggplotly(p11)
```